---
title: "Expected Values"
author: "MATH 345"
date: "Fall 2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This note shows how to create R functions to calculate expected values of random variables (r.v.) and functions thereof.

## 1. Function definitions

**Function to calculate the expected value of a random variable**

Inputs: `y`: possible values of the r.v.; `py`: associated probabilities $p(y) = P(Y = y)$.

Output: Expected value $E(Y) = \sum_{y} y p(y)$

```{r, expected value}
E <- function(y, py)
{
	stopifnot(length(y) == length(py))
	out <- sum(y * py)
	out
}
```

Note: you can use the command `return` at the end of a function to tell R to return a value (for example, `return(out` in the function above) but it's not necessary. If the last line is just the name of an object, R will know to return it.

**Function to calculate the variance of a random variable**

Inputs: `y`: possible values of the r.v.; `py`: associated probabilities $p(y) = P(Y = y)$.

Output: Variance $V(Y) = \sum_{y} ((y-\mu)^2 p(y))$ where $\mu = E(Y)$.

```{r}
V <- function(y, py) {
	mu <- E(y, py)
	sum((y-mu)^2 * py)
}
```

**Function to calculate the expected value of a function of a random variable**

Inputs: `y`: possible values of the r.v.; `py`: associated probabilities $p(y) = P(Y = y)$; `g`: function (should be able to take vector inputs).

Output: $E(g(Y)) = \sum_{y} g(y) p(y)$

```{r, LOTUS}
EgY <- function(y, py, g) sum(g(y) * py)
```

-   Note: if a function definition holds in one line, there's no need to put squiggly brackets {} around it
-   Law Of The Unconscious Statistician (LOTUS): <https://en.wikipedia.org/wiki/Law_of_the_unconscious_statistician>

## 2. Example: Textbook exercise 3.12

Probability table:

```{r}
p <- c(.4, .3, .2, .1)
y <- 1:4
```

Expected value of $Y$:
```{r}
E(y, p) 
```
Note: as in other programming languages, when you *call* a function on some R variables (here `y` and `p`), these variables do not need to have the same names as those used to *define* the function (here `y` and `py`).

Expected value of inverse of Y ($E(1/Y)$):

```{r}
g1 <- function(y) 1/y
EgY(y, p, g1)  
```

Expected value $E(Y^2 - 1)$:
```{r}
g2 <- function(y) y^2 - 1
EgY(y, p, g2)
```

Variance of $Y$ ($V(Y)$):
```{r}
V(y, p) 
```
